---
title: "Swiss Cantonal 89-18 Cross Basis Selection"
author: "Christoph Kestenholz"
date: "18 May 2021"
output: html_document
editor_options:
  chunk_output_type: console
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"oeschger_logo_rgb.jpg\" style=\"float: upper;width: 170px;\"/>')
   });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using a basic model, we test different CB-humidity specifications and select the 
one with lowest grand q-AIC. We use the basic temperature cb and the model for summer months.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# LOAD LIBRARIES
library(dlnm) ; library(splines) ; library(tsModel) ; library(lubridate)
library(mixmeta); library(ggplot2);
library("Epi") ; library("splines")

# SPECIFY IF WORKING FROM DESKTOP OR LAPTOP
# dirpre <- "C:/Users/ppehavic/OD 1/LSHTM/MCC/MCCOzone/Analysis/"
setwd("D:/data/TempHumidity/02")
dirpre <- "D:/data/TempHumidity/"


# LOAD DATA (CLEAN MCCOZONE)
# load(paste(dirpre,"output/swiss_cantonal_89-18.rds", sep=""))
dlist <- readRDS(paste(dirpre,"00/DataTempHumidity_CK.rds", sep=""))

#dlist <- readRDS("swiss_cantonal_89-18.rds")

# FUNCTION FOR COMPUTING THE Q-AIC
QAIC <- function(model) {
  phi <- summary(model)$dispersion
	loglik <- sum(dpois(model$y, model$fitted.values, log=TRUE))
  return(-2*loglik + 2*summary(model)$df[3]*phi)
}

```

## 1. PARAMETERS - CB SELECTION OF HUMIDITY

### DEFINE DIFFERENT SPECIFICATIONS FOR EACH DIMENSION OF THE CB

NOTE: We keep constant the definition of cb-temperature and change the parameters for cb-humidity
Create one matrix to define the alternative combinations of the two functions.

```{r echo=TRUE}
# CREATE TWO MATRIX WITH THE DIFFERENT COMB OF FUNCTIONS AND 
comb <- expand.grid(1:4,1:5)
```

### DEFINE THE OTHER TERMS OF THE MODEL

```{r, echo=TRUE}
# DEGREE OF FREEDOM FOR SEASONALITY
dfseas <- 4

# DEGREE OF FREEDOM FOR TREND
dftrend <- 1

# MODEL FORMULA
formula1 <- deaths ~ cbt + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)) # Max Temp
formula2 <- deaths ~ cbh + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)) # Partial Humidity Pressue
formula3 <- deaths ~ cbt + cbh + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)) # Max Temp + Partial Humidity Pressure
formula4 <- deaths ~ cbrelh + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)) # Relative Humidity
formula5 <- deaths ~ cbt + cbrelh + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)) # Max Temp + Relative Humidity
```

For each combination of functions, run the location-specific model and store
the estimates, objects and qaic.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
# CREATE AN EMPTY MATRIX TO STORE THE QAIC
qaic_tabPP <- array(NA,dim=c(length(dlist),nrow(comb), 3)
    ,dimnames=list(names(dlist),c(paste0("comb",1:nrow(comb))),
                   c("temp_null","hum_null","comb")))

qaic_tabRH <- array(NA,dim=c(length(dlist),nrow(comb), 3)
    ,dimnames=list(names(dlist),c(paste0("comb",1:nrow(comb))),
                   c("temp_null","hum_null","comb")))

crosspredall_humnull_list <- list()
crosspredall_tempnull_list <- list()
crosspredall_humcomb_list <- list()
crosspredall_tempcomb_list <- list()
crosspredall_humrelnull_list <- list()
crosspredall_humrelcomb_list <- list()
crosspredall_temprelcomb_list <- list()

# RUN THE MODEL FOR EACH COMBINATION OF FUNCTIONS
for (j in 1:nrow(comb)){
 
  cat(j,"")

  crosspredall_humnull <- list()
  crosspredall_tempnull <- list()
  crosspredall_humcomb <- list()
  crosspredall_tempcomb <- list()
  crosspredall_humrelnull <- list()
  crosspredall_humrelcomb <- list()
  crosspredall_temprelcomb <- list()
  
 # RUN THE LOOP ACROSS CANTONS
 for (i in c(1,12,21,23,25)){
 #for(i in seq(length(dlist))) {

  # PRINT
  cat(i,"")

  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)

  #DEFINE CROSSBASIS TEMPERATURE
  cbt <- crossbasis(data$`max temperature [deg]`, lag=10,
   argvar=list(fun="bs",knots=quantile(data$`max temperature [deg]`,c(50,90)/100,na.rm=T), degree=2),
   arglag=list(fun="ns", knots=logknots(10,2)), group=as.factor(data$year)) # could be 7days according to previous formula?

  # DEFINE CROSSBASIS HUMIDITY
  # DIFFERENT FUNCTIONS OF THE LAG DIMENSION
  maxlag <- 5 
  arglag1o <- list(knots=logknots(maxlag,1))
  arglag2o <- list(fun="strata", breaks=1)
  arglag3o <- list(fun="integer")
  arglag4o <- list(fun="ns", knots=logknots(maxlag,1))
  arglag5o <- list(fun="strata", breaks=c(1,3))

  # DEFINE DIFFERENT FUNCITONS OF THE VAR DIMENSION - PARTIAL PRESSURE
  argvar1o <- list(fun="lin")
  argvar2o <- list(fun="ns", knots=quantile(data$`partial pressure [hPa]`, c(0.5, 0.9), na.rm=T),
                 Boundary.knots=range(data$`partial pressure [hPa]`, na.rm=T))
  argvar3o <- list(fun="ns", knots=quantile(data$`partial pressure [hPa]`, c(0.75), na.rm=T),
                 Boundary.knots=range(data$`partial pressure [hPa]`, na.rm=T))
  argvar4o <- list(fun="thr", side="h", thr.value=quantile(data$`partial pressure [hPa]`, 0.9, na.rm=T))
  
  cbh <- crossbasis(data$`partial pressure [hPa]`, lag=5,
   argvar=get(paste("argvar",comb[j,1],"o", sep="")),
   arglag=get(paste("arglag",comb[j,2],"o", sep="")),  group=as.factor(data$year))
  
  # DEFINE DIFFERENT FUNCITONS OF THE VAR DIMENSION - REL HUMIDITY
  argvar1o <- list(fun="lin")
  argvar2o <- list(fun="ns", knots=quantile(data$`relative humidity [%]`, c(0.5, 0.9), na.rm=T),
                 Boundary.knots=range(data$`relative humidity [%]`, na.rm=T))
  argvar3o <- list(fun="ns", knots=quantile(data$`relative humidity [%]`, c(0.75), na.rm=T),
                 Boundary.knots=range(data$`relative humidity [%]`, na.rm=T))
  argvar4o <- list(fun="thr", side="h", thr.value=quantile(data$`relative humidity [%]`, 0.9, na.rm=T))
  
  cbrelh <- crossbasis(data$`relative humidity [%]`, lag=5,
   argvar=get(paste("argvar",comb[j,1],"o", sep="")),
   arglag=get(paste("arglag",comb[j,2],"o", sep="")),  group=as.factor(data$year))
  

  # CREATE YDAY
  data$yday <- yday(data$date)
  
  # MODEL ONLY temperature 
  model <- glm(formula1,data,family=quasipoisson,na.action="na.exclude")
  qaic_tabPP[i,j,1] <- QAIC(model)
  crosspredall_tempnull[[i]] <- crosspred(cbt,model,cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 
  
  # MODEL ONLY humidity PP
  model <- glm(formula2,data,family=quasipoisson,na.action="na.exclude")
  qaic_tabPP[i,j,2] <- QAIC(model)
  crosspredall_humnull[[i]] <- crosspred(cbh,model,cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 by=0.1) 
  
  # MODEL temp +  humidity PP
  model <- glm(formula3,data,family=quasipoisson,na.action="na.exclude")
  qaic_tabPP[i,j,3] <- QAIC(model)
  crosspredall_humcomb[[i]] <- crosspred(cbh,model,cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 by=0.1) 
  crosspredall_tempcomb[[i]] <- crosspred(cbt,model,cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 
  
  # MODEL ONLY humidity RH
  model <- glm(formula4,data,family=quasipoisson,na.action="na.exclude")
  qaic_tabRH[i,j,2] <- QAIC(model)
  crosspredall_humrelnull[[i]] <- crosspred(cbrelh,model,cen=mean(data$`relative humidity [%]`,na.rm=T),
                                 by=0.1) 
  
  # MODEL temp +  humidity RH
  model <- glm(formula5,data,family=quasipoisson,na.action="na.exclude")
  qaic_tabRH[i,j,3] <- QAIC(model)
  crosspredall_humrelcomb[[i]] <- crosspred(cbrelh,model,cen=mean(data$`relative humidity [%]`,na.rm=T),
                                 by=0.1) 
  crosspredall_temprelcomb[[i]] <- crosspred(cbt,model,cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 

 }
  
crosspredall_humnull_list[[j]] <- crosspredall_humnull
crosspredall_tempnull_list[[j]] <- crosspredall_tempnull
crosspredall_humcomb_list[[j]] <- crosspredall_humcomb
crosspredall_tempcomb_list[[j]] <- crosspredall_tempcomb
crosspredall_humrelnull_list[[j]] <- crosspredall_humrelnull
crosspredall_humrelcomb_list[[j]] <- crosspredall_humrelcomb
crosspredall_temprelcomb_list[[j]] <- crosspredall_temprelcomb
}

# summaryqaicPP <- apply(qaic_tabPP,2:3,sum) # for all 26 cantons
summaryqaicPP <- apply(qaic_tabPP[c(1,12,21,23,25),,],2:3,sum)

# summaryqaicRH <- apply(qaic_tabRH,2:3,sum) # for all 26 cantons
summaryqaicRH <- apply(qaic_tabRH[c(1,12,21,23,25),,],2:3,sum)


combPP <- which.min(summaryqaicPP[,3]) # linear & strata

#combRH <- which.min(summaryqaicRH[,3]) # linear & strata
combRH <- combPP

# SAVE THE WORKSPACE
save.image("output/cbselection.RData")

# # LOAD STORED RESULTS 
# load("output/cbselection.RData")

plot(1:20,summaryqaicPP[,2], main="CB selection HUMI_NULL")
plot(1:20,summaryqaicPP[,3], main="CB selection COMB")

# plot(1:20,summaryqaicRH[,2], main="CB selection RH HUMI_NULL") # relative humidity
# plot(1:20,summaryqaicRH[,3], main="CB selection RH COMB") # relative humidity

which.min(summaryqaicPP[,3]) # linear & strata
# which.min(summaryqaicRH[,3]) # # relative humidity: linear & strata


```

## 2. SELCTED MODEL - MAIN RESULTS

Plot the results for temperature and humidity for the 5 cantons selected
```{r}
# Save form
cptempnull <- cphumppnull <- cphumppcomb <- 
  cptempcombpp <- cphumrhnull <- cphumrhcomb <- 
  cptempcombrh <-  cptempnulllag <- cphumppnulllag <-
  cphumppcomblag <- cptempcombpplag <- cphumrhnulllag <-
  cphumrhcomblag <- cptempcombrhlag <- list()

selcantons <- c(1,12,21,23,25)

for(i in seq(length(dlist))) {

  # PRINT
  cat(i,"")

  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)

  #DEFINE CROSSBASIS TEMPERATURE
  cbt <- crossbasis(data$`max temperature [deg]`, lag=10,
   argvar=list(fun="bs",knots=quantile(data$`max temperature [deg]`,c(50,90)/100,na.rm=T), degree=2),
   arglag=list(fun="ns", knots=logknots(10,2)), group=as.factor(data$year)) # could be 7days according to previous formula?

  # DEFINE CROSSBASIS HUMIDITY
  # DIFFERENT FUNCTIONS OF THE LAG DIMENSION
  maxlag <- 5 
  arglag2o <- list(fun="strata", breaks=1)

  
  # DEFINE DIFFERENT FUNCITONS OF THE VAR DIMENSION - PARTIAL PRESSURE
  argvar1o <- list(fun="lin")
  
  cbh <- crossbasis(data$`partial pressure [hPa]`, lag=maxlag,
   argvar=argvar1o,
   arglag=arglag2o,  group=as.factor(data$year))

  cbrelh <- crossbasis(data$`relative humidity [%]`, lag=maxlag,
   argvar=argvar1o,
   arglag=arglag2o,  group=as.factor(data$year))
  
  # CREATE YDAY
  data$yday <- yday(data$date)
  
  # MODEL ONLY temperature 
  model <- glm(formula1,data,family=quasipoisson,na.action="na.exclude")
  cptempnull[[i]] <- crossreduce(cbt,model,
                                 cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 
  cptempnulllag[[i]] <- crossreduce(cbt,model,type="var",
                                 cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 value=quantile(data$`max temperature [deg]`,0.99, na.rm=T),
                                 by=0.1) 
  
  # MODEL ONLY humidity PP
  model <- glm(formula2,data,family=quasipoisson,na.action="na.exclude")
  cphumppnull[[i]] <- crossreduce(cbh,model,cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 by=0.1) 
  cphumppnulllag[[i]] <- crossreduce(cbh,model,type="var",
                                 cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 value=quantile(data$`partial pressure [hPa]`,0.99, na.rm=T),
                                 by=0.1) 
  
  # MODEL temp +  humidity PP
  model <- glm(formula3,data,family=quasipoisson,na.action="na.exclude")
  cphumppcomb[[i]] <- crossreduce(cbh,model,cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 by=0.1) 
  cptempcombpp[[i]] <- crossreduce(cbt,model,cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 
  
  cphumppcomblag[[i]] <- crossreduce(cbh,model,type="var",
                                 cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 value=quantile(data$`partial pressure [hPa]`,0.99, na.rm=T),
                                 by=0.1) 
  cptempcombpplag[[i]] <- crossreduce(cbt,model,type="var",
                                 cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 value=quantile(data$`max temperature [deg]`,0.99, na.rm=T),
                                 by=0.1) 
  
  # MODEL ONLY humidity RH
  model <- glm(formula4,data,family=quasipoisson,na.action="na.exclude")
  cphumrhnull[[i]] <- crossreduce(cbrelh,model,cen=mean(data$`relative humidity [%]`,na.rm=T),
                                 by=0.1) 
  cphumrhnulllag[[i]] <- crossreduce(cbrelh,model,type="var",
                                 cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 value=quantile(data$`partial pressure [hPa]`,0.99, na.rm=T),
                                 by=0.1) 
  
  # MODEL temp +  humidity RH
  model <- glm(formula5,data,family=quasipoisson,na.action="na.exclude")
  cphumrhcomb[[i]] <- crossreduce(cbrelh,model,cen=mean(data$`relative humidity [%]`,na.rm=T),
                                 by=0.1) 
  cptempcombrh[[i]] <- crossreduce(cbt,model,cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 by=0.1) 
  
  cphumrhcomblag[[i]] <- crossreduce(cbrelh,model,type="var",
                                 cen=mean(data$`partial pressure [hPa]`,na.rm=T),
                                 value=quantile(data$`partial pressure [hPa]`,0.99, na.rm=T),
                                 by=0.1) 
  cptempcombrhlag[[i]] <- crossreduce(cbt,model,type="var",
                                 cen=mean(data$`max temperature [deg]`,na.rm=T),
                                 value=quantile(data$`max temperature [deg]`,0.99, na.rm=T),
                                 by=0.1) 

}


## PLOTS 

listcantons <- names(dlist)

# 1
png(filename = "er_humpptemp_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,4))

for (i in selcantons){
  plot(cphumppnull[[i]], main=paste0("Humidity (null) ",
                listcantons[i]), ylab="RR", xlab="Partial pressure")
  plot(cptempnull[[i]], main=paste0("Temperature (null) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
  plot(cphumppcomb[[i]], main=paste0("Humidity (comb) ",
                listcantons[i]), ylab="RR", xlab="Partial pressure")
  plot(cptempcombpp[[i]],  main=paste0("Temperature (comb) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
}
dev.off()

# 2
png(filename = "er_humreltemp_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,4))

for (i in selcantons){
  plot(cphumrhnull[[i]],main=paste0("Humidity (null) ",
                listcantons[i]), ylab="RR", xlab="Rel Hum")
  plot(cptempnull[[i]], main=paste0("Temperature (null) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
  plot(cphumrhcomb[[i]], main=paste0("Humidity (comb) ",
                listcantons[i]), ylab="RR", xlab="Rel Hum")
  plot(cptempcombrh[[i]], main=paste0("Temperature (comb) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
}
dev.off()

# 3
png(filename = "lag_humpptemp_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,4))

for (i in selcantons){
  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)
  
  plot(cphumppnulllag[[i]],
       main=paste0("Humidity (null) ", listcantons[i]), 
       ylab="RR", xlab="Partial pressure")
  plot(cptempnulllag[[i]], 
       main=paste0("Temperature (null) ", listcantons[i]),
       ylab="RR", xlab="Temperature")
  plot(cphumppcomblag[[i]],  
       main=paste0("Humidity (comb) ", listcantons[i]), 
       ylab="RR", xlab="Partial pressure")
  plot(cptempcombpplag[[i]],  
       main=paste0("Temperature (comb) ", listcantons[i]),
       ylab="RR", xlab="Temperature")
}
dev.off()

# 4
png(filename = "lag_humrhtemp_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,4))

for (i in selcantons){
  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)
  
  plot(cphumrhnulllag[[i]],
       main=paste0("Humidity (null) ", listcantons[i]), 
       ylab="RR", xlab="Rel humidity")
  plot(cptempnulllag[[i]], 
       main=paste0("Temperature (null) ", listcantons[i]),
       ylab="RR", xlab="Temperature")
  plot(cphumrhcomblag[[i]],  
       main=paste0("Humidity (comb) ", listcantons[i]), 
       ylab="RR", xlab="Rel humidity")
  plot(cptempcombrhlag[[i]],  
       main=paste0("Temperature (comb) ", listcantons[i]),
       ylab="RR", xlab="Temperature")
}
dev.off()


# COMBINED PLOTS

colci <- do.call(rgb,c(as.list(col2rgb("red")),alpha=255/8,max=255))

png(filename = "er_humtempppcomb_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,2))
for (i in selcantons){
  plot(cphumppnull[[i]],  main=paste0("Humidity (null) ",
                listcantons[i]), ylab="RR", xlab="Partial pressure")
  lines(cphumppcomb[[i]],type="l",ci="area",  col="red", ci.arg=list(col=colci))
  legend("top", legend=c("null", "comb"), col=c("black", "red"), lty=1, cex=0.8)
  
  plot(cptempnull[[i]],main=paste0("Temperature (null) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
  lines(cptempcombpp[[i]],type="l",ci="area",  col="red", ci.arg=list(col=colci) )
  legend("top", legend=c("null", "comb"), col=c("black", "red"), lty=1, cex=0.8)
}
dev.off()


png(filename = "er_humtemprhcomb_5cantons.png", 
    width = 2000, height = 2000, units = "px", pointsize = 24,
    bg = "white",  res = NA)

par(mfrow=c(5,2))
for (i in selcantons){
  plot(cphumrhnull[[i]],  main=paste0("Humidity (null) ",
                listcantons[i]), ylab="RR", xlab="Rel Humidity")
  lines(cphumrhcomb[[i]],type="l",ci="area",  col="red", ci.arg=list(col=colci))
  legend("top", legend=c("null", "comb"), col=c("black", "red"), lty=1, cex=0.8)
  
  plot(cptempnull[[i]],main=paste0("Temperature (null) ",
                listcantons[i]), ylab="RR", xlab="Temperature")
  lines(cptempcombpp[[i]],type="l",ci="area",  col="red", ci.arg=list(col=colci) )
  legend("top", legend=c("null", "comb"), col=c("black", "red"), lty=1, cex=0.8)
}
dev.off()


```


## 3. INTERACTION MODELS

```{r, echo=TRUE}
 ########################
 # Interaction models  1 ###
 ########################

# EFFECT OF TEMPERATURE DURING HUMID/NO HUMID DAYS
 
 # CREATE THE OBJECTS TO STORE THE RESULTS
 redall.n <- redall.1 <- redall.2 <- redall.int <-list()
 
 # RUN THE LOOP ACROSS CANTONS 
 
 for(i in seq(length(dlist))) {

  # PRINT
  cat(i,"")

  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)
  
  data$yday <- as.numeric(format(data$date,"%j"))

  #DEFINE CROSSBASIS TEMPERATURE and define DUMMY-terms
  cbt <- crossbasis(data$`max temperature [deg]`, lag=10,
   argvar=list(fun="bs",knots=quantile(data$`max temperature [deg]`,c(50,90)/100,na.rm=T), degree=2),
   arglag=list(fun="ns", knots=logknots(10,2)), group=as.factor(data$year)) # could be 7days according to previous formula?
  
  # CREATE DUMMY VARIABLE FOR HUMIDITY
  data$dummy1<- ifelse(data$`partial pressure [hPa]` > quantile(data$`partial pressure [hPa]`,0.9,na.rm=T),1,0) # very humid
  data$dummy2 <- ifelse(data$`partial pressure [hPa]` <= quantile(data$`partial pressure [hPa]`,0.9,na.rm=T),1,0) # dry

  # CREATE INTERACTION TERMS
  int1 <- data$dummy1*cbt # very humid
  int2 <- data$dummy2*cbt # dry

  modnull <- glm(deaths ~ cbt + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data)

  modint1 <- glm(deaths ~ cbt + int1 + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data) 

  modint2 <- glm(deaths ~ cbt + int2 + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data)
  
  # PREDICT
  redtemp.n <- crossreduce(cbt,modnull,cen=10, by=0.1, 
                           from=round(quantile(data$`max temperature [deg]`, 0.25, na.rm=T),1),
                           to=round(quantile(data$`max temperature [deg]`, 0.98, na.rm=T),1) ) 
  redtemp.1 <- crossreduce(cbt,modint1,cen=10, by=0.1, 
                           from=round(quantile(data$`max temperature [deg]`, 0.25, na.rm=T),1),
                           to=round(quantile(data$`max temperature [deg]`, 0.98, na.rm=T),1) ) 
  redtemp.2 <- crossreduce(cbt,modint2,cen=10, by=0.1, 
                           from=round(quantile(data$`max temperature [deg]`, 0.25, na.rm=T),1),
                           to=round(quantile(data$`max temperature [deg]`, 0.98, na.rm=T),1) ) 

  # RESTRCIT SEARCH MMT BETWEEN 25:98 PCT
  mmh.n <- as.numeric(names(which.min(redtemp.n$fit)))
  mmh.1 <- as.numeric(names(which.min(redtemp.1$fit)))
  mmh.2 <- as.numeric(names(which.min(redtemp.2$fit)))
  
  redall.n[[i]] <- crossreduce(cbt,modnull, cen=mmh.n, by=0.2)  
  redall.1[[i]] <- crossreduce(cbt,modint1, cen=mmh.1, by=0.2)  # temp effect when int1==0 (dry)
  redall.2[[i]] <- crossreduce(cbt,modint2, cen=mmh.2, by=0.2)  # temp effect when int2==0 (very humid)
  redall.int[[i]] <- crossreduce(int1,modint1, cen=mmh.1, by=0.2)  
 }
 
 # set plot settings for ci:
 col1 <- do.call(rgb, c(as.list(col2rgb("orange")), alpha=255/15, max=255)) # mixed effect
 col2 <- do.call(rgb, c(as.list(col2rgb("blue")), alpha=255/10, max=255)) # humid days
 col3 <- do.call(rgb, c(as.list(col2rgb("red")), alpha=255/10, max=255)) # dry days
 col4 <- do.call(rgb, c(as.list(col2rgb("forestgreen")), alpha=255/10, max=255)) # interaction

 
 png(filename = "intmodel_temp_5cantons.png", 
    width = 2000, height = 1500, units = "px", pointsize = 24,
    bg = "white",  res = NA)
 
 par(mfrow=c(2,3))

 for (k in c(1,12,21,23,25)){
 plot(redall.n[[k]], col="orange", ci.arg=list(col=col1), 
      main=paste0(listcantons[k], " interaction model", sep="  "), xlab="mean temperature (\u00B0C)", 
       ylab="RR", ylim=c(0.75,2.0))
 lines(redall.1[[k]], col="red", ci="area", ci.arg=list(col=col3))
 lines(redall.2[[k]], col="blue", ci="area", ci.arg=list(col=col2),)
 lines(redall.int[[k]], col="forestgreen", ci="area", ci.arg=list(col=col4))

 abline(v= quantile(data$`max temperature [deg]`, 0.99, na.rm=T),col = "gray60", lty = 3)
 # legend("top", legend=c("modelnull", "dry", "humid", "interaction"),
 #       col=c("orange","red", "blue",  "green"), lty=1:4, cex=0.8)
 }
 plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1, main="Legend")
 legend("top", legend=c("null", "mild humidity", "high humidity", "interaction"),
        col=c("orange","red", "blue",  "forestgreen"), lty=1, cex=1.5)

 
dev.off()
 

# EFFECT OF HUMIDITY DURING HOT/NO HOT DAYS
 
 # CREATE THE OBJECTS TO STORE THE RESULTS
 redall.n <- redall.1 <- redall.2 <- redall.int <-list()
 
 # RUN THE LOOP ACROSS CANTONS 
 
 for(i in seq(length(dlist))) {

  # PRINT
  cat(i,"")

  # EXTRACT THE DATA
  data <- dlist[[i]]

  # SUMMER PERIOD
  data <- subset(data, month %in% 5:9)
  
  data$yday <- as.numeric(format(data$date,"%j"))

  #DEFINE CROSSBASIS TEMPERATURE and define DUMMY-terms
  cbt <- crossbasis(data$`max temperature [deg]`, lag=10,
   argvar=list(fun="bs",knots=quantile(data$`max temperature [deg]`,c(50,90)/100,na.rm=T), degree=2),
   arglag=list(fun="ns", knots=logknots(10,2)),group=as.factor(data$year)) # could be 7days according to previous formula?
  
  cbh <- crossbasis(data$`partial pressure [hPa]`, lag=5,
   argvar=list(fun="lin"),
   arglag=list(fun="strata", breaks=1),  group=as.factor(data$year))
  
  # CREATE DUMMY VARIABLE FOR TEMPERATURE
  data$dummy1<- ifelse(data$`max temperature [deg]` > quantile(data$`max temperature [deg]`,0.9,na.rm=T),1,0) # very humid
  data$dummy2 <- ifelse(data$`max temperature [deg]` <= quantile(data$`max temperature [deg]`,0.9,na.rm=T),1,0) # dry

  # CREATE INTERACTION TERMS
  int1 <- data$dummy1*cbh # very humid
  int2 <- data$dummy2*cbh # dry

  modnull <- glm(deaths ~ cbt + cbh + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data)

  modint1 <- glm(deaths ~ cbt + cbh + int1 + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data) 

  modint2 <- glm(deaths ~ cbt + cbh + int2 + dow + ns(yday,df=dfseas):factor(year) + ns(date,df=round(length(unique(year))/dftrend/10)), 
  family=quasipoisson(), data)
  
  # PREDICT
  redall.n[[i]] <- crossreduce(cbh,modnull, cen=quantile(data$`partial pressure [hPa]`, 0.25, na.rm=T), by=0.2)  
  redall.1[[i]] <- crossreduce(cbh,modint1, cen=quantile(data$`partial pressure [hPa]`, 0.25, na.rm=T), by=0.2)  
  redall.2[[i]] <- crossreduce(cbh,modint2, cen=quantile(data$`partial pressure [hPa]`, 0.25, na.rm=T), by=0.2)  
  redall.int[[i]] <- crossreduce(int1,modint1, cen=quantile(data$`partial pressure [hPa]`, 0.25, na.rm=T), by=0.2)  
 }
 

 # set plot settings for ci:
 col1 <- do.call(rgb, c(as.list(col2rgb("orange")), alpha=255/15, max=255)) # mixed effect
 col2 <- do.call(rgb, c(as.list(col2rgb("blue")), alpha=255/10, max=255)) # hot days
 col3 <- do.call(rgb, c(as.list(col2rgb("red")), alpha=255/10, max=255)) # no hot days
 col4 <- do.call(rgb, c(as.list(col2rgb("forestgreen")), alpha=255/10, max=255)) # interaction

 
png(filename = "intmodel_humpp_5cantons.png", 
    width = 2000, height = 1500, units = "px", pointsize = 24,
    bg = "white",  res = NA)
  
 par(mfrow=c(2,3))

 for (k in c(1,12,21,23,25)){
 plot(redall.n[[k]], col="orange", ci.arg=list(col=col1), 
      main=paste0(listcantons[k], " canton interaction models", sep="  "), 
      xlab="Partial pressure", 
       ylab="RR", ylim=c(0.80,1.2))
 lines(redall.1[[k]], col="red", ci="area", ci.arg=list(col=col3))
 lines(redall.2[[k]], col="blue", ci="area", ci.arg=list(col=col2),)
 lines(redall.int[[k]], col="forestgreen", ci="area", ci.arg=list(col=col4))

 abline(v= quantile(data$`max temperature [deg]`, 0.99, na.rm=T),col = "gray60", lty = 3)
 # legend("top", legend=c("modelnull", "dry", "humid", "interaction"),
 #       col=c("orange","red", "blue",  "forestgreen"), lty=1:4, cex=0.8)
 } 
 plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1, main="Legend")
 legend("top", legend=c("null", "mild temperature", "high temperature", "interaction"),
        col=c("orange","red", "blue",  "forestgreen"), lty=1, cex=1.5)

 dev.off()
 
```
## 4. RESULTS AT THE 99P IN TABLE

```{r, echo=TRUE}
# summarize results:
res_TempHumidity <- matrix(NA, nrow=26, ncol=7)
colnames(res_TempHumidity) <- c("cptempnull", "cptempcombpp", "cptempcombrh", "cphumppnull", "cphumppcomb",
                                "cphumrhnull", "cphumrhcomb")
#rownames(res_TempHumidity) <- c("Zurich", "Basel-Stadt", "Ticino", "Valais", "Geneva")
rownames(res_TempHumidity) <- names(dlist)

 for (k in c(1,12,21,23,25)){
   
   # load the summer data
   # EXTRACT THE DATA
   data <- dlist[[k]]
   data <- subset(data, month %in% 5:9)
   # calculate the 99th percentile of the variable
   NNPC <- round(quantile(data$`max temperature [deg]`, 0.99, na.rm=T), digits=1)
   
   res_TempHumidity[k,1] <- paste0(round(cptempnull[[k]]$RRfit[as.character(NNPC)],2), " [", round(cptempnull[[k]]$RRlow[as.character(NNPC)],2), " - ", round(cptempnull[[k]]$RRhigh[as.character(NNPC)],2) , "]")
   
   res_TempHumidity[k,2] <- paste0(round(cptempcombpp[[k]]$RRfit[as.character(NNPC)],2), " [", round(cptempcombpp[[k]]$RRlow[as.character(NNPC)],2), " - ", round(cptempcombpp[[k]]$RRhigh[as.character(NNPC)],2) , "]")
   
   res_TempHumidity[k,3] <- paste0(round(cptempcombrh[[k]]$RRfit[as.character(NNPC)],2), " [", round(cptempcombrh[[k]]$RRlow[as.character(NNPC)],2), " - ", round(cptempcombrh[[k]]$RRhigh[as.character(NNPC)],2) , "]")
   
   res_TempHumidity[k,4] <- paste0(round(cphumppnull[[k]]$coefficients[[1]], digits=4), " per hPa of PP") # linear
   
   res_TempHumidity[k,5] <- paste0(round(mean(cphumppcomb[[k]]$coefficients[[1]]), digits=4), " per hPa of PP") # linear 
   
   res_TempHumidity[k,6] <- paste0(round(mean(cphumrhnull[[k]]$coefficients[[1]]), digits=4), " per % of RH") # linear 
   
   res_TempHumidity[k,7] <- paste0(round(mean(cphumrhcomb[[k]]$coefficients[[1]]), digits=4), " per % of RH") # linear 
   
 }

write.table(res_TempHumidity[c(1,12,21,23,25), ], file = "res_TempHumidity_99p.txt", sep = ",", quote = FALSE, row.names = T)


```